%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%% Bachelor Thesis Template of the PRG Group
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt,twoside,a4paper,fleqn,bibliography=totocnumbered]{report}

% some often used and useful packages (others can be added if necessary)
\usepackage{latexsym}
\usepackage[utf8]{inputenc}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pifont}
\usepackage{psfrag}
\usepackage{epsfig}
\usepackage{bm}
\usepackage{wasysym}
\usepackage[table]{xcolor}
\usepackage{setspace}
\usepackage{framed}
\usepackage{float,pdflscape}
\usepackage{url}
\usepackage{color}
\usepackage{minted}
\usepackage{algpseudocode}
\usepackage[margin=1.1in,bindingoffset=.2in]{geometry} 
\usepackage{longtable}
\usepackage{tabularx}   % For better table formatting
\usepackage{multicol}    % For multi-column layouts

% Set global table parameters for more compact tables
\renewcommand{\arraystretch}{0.95}  % More compact rows
\setlength{\tabcolsep}{4pt}         % Tighter column spacing

% Add consistent notation macros for vectors, matrices, and number sets
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\C}{\mathbb{C}}

\onehalfspacing



\begin{document}


% frontmatter
\input title.tex
\pagenumbering{roman}
\setcounter{page}{1}
\cleardoublepage
\include{abstract}
%\cleardoublepage
%\input acknowledgments.tex % delete this and next line if not applicable
%\cleardoublepage
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}
\setcounter{page}{1}

% The major parts 
\include{introduction}
\include{theory}
\include{method}
\include{experiments}
\include{conclusions}

% endmatter
\include{appendix}
\cleardoublepage
\addcontentsline{toc}{chapter}{Bibliography}
% Comment out BibTeX commands
%\bibliography{../references}
%\bibliographystyle{unsrt}    	             

% Manual bibliography section
\begin{thebibliography}{1}

%Estimating algebraic connectivity
\bibitem{estalgeb1} A. Bertrand and M. Moonen, ``Seeing the bigger picture: How nodes can learn their place within a complex ad hoc network topology'' \emph{IEEE Signal Process. Mag.}, vol. 30, pp. 71-82, 2013.

%Leiden algorithm
\bibitem{leiden} V. A. Traag, L. Waltman, and N. J. van Eck, ``From Louvain to Leiden: guaranteeing well-connected communities,'' \emph{Scientific Reports}, vol. 9, no. 1, pp. 5233, 2019.

% KNN Outlier Detection
\bibitem{knn_outlier} Z. Qi, J. Zhang, X. Chen, and X. Qi, ``Comparative Study of Neighbor-based Methods for Local Outlier Detection,'' \emph{arXiv preprint arXiv:2405.19247}, 2024.

\bibitem{iztech_info} Izmir Institute of Technology, "Kurumsal Bilgiler," [Online]. Available: https://uio.iyte.edu.tr/kurumsal-bilgiler/

% Campus transit planning
\bibitem{dell2016campus} L. Dell'Olio, K. Bertolaccini, and A. Ibeas, ``Campus transit for the twenty-first century: an object-oriented model for the UCSC loop,'' \emph{European Journal of Operational Research}, vol. 248, no. 1, pp. 143-154, 2016.

% Vehicle routing problems
\bibitem{toth2014vehicle} P. Toth and D. Vigo, ``Vehicle routing: problems, methods, and applications,'' \emph{Society for Industrial and Applied Mathematics}, 2014.

% Smart campus transit planning
\bibitem{kaviani2019smart} H. Kaviani, M. M. Karbasioun, M. K. O'Brien, O. Salamah, K. M. Kelley, and B. Krishnamachari, ``Smart campus transit system planning and management with multi-agent simulation,'' \emph{Sustainability}, vol. 11, no. 24, pp. 7031, 2019.

% Multi-depot vehicle routing
\bibitem{saberi2017models} M. Saberi and H. S. Mahmassani, ``Models and algorithms for the capacitated pre-scheduled multi depot multiple traveling salesman problem with heterogeneous vehicles,'' \emph{Transportation Research Part C: Emerging Technologies}, vol. 83, pp. 263-279, 2017.

% Sustainable mobility on university campuses
\bibitem{guido2017sustainable} G. Guido, A. Sestino, V. Marín-Puchades, M. J. Rodríguez-Fórtiz, S. R. Faria, A. Temporelli, J. Massiani, and F. Milicchio, ``Sustainable mobility at a university campus: walking preferences and the use of electric minibus,'' \emph{International Journal of Sustainable Transportation}, vol. 12, no. 10, pp. 1-10, 2017.

% Outlier detection in transportation data
\bibitem{lu2020outlier} S. Lu, J. Cui, X. Lian, Z. Fan, Z. Zheng, and Y. Tong, ``Outlier detection for transportation big data: Lessons from implementations in production,'' \emph{Proceedings of the ACM on Management of Data}, vol. 1, no. 1, pp. 1-29, 2020.

% Graph-based transportation network analysis
\bibitem{tarapata2019graph} Z. Tarapata, ``Graph-based approaches to transportation networks analysis,'' \emph{Journal of Telecommunications and Information Technology}, vol. 3, pp. 51-63, 2019.

% Data-driven transportation systems
\bibitem{zhang2018data} J. Zhang, F.-Y. Wang, K. Wang, W.-H. Lin, X. Xu, and C. Chen, ``Data-driven intelligent transportation systems: A survey,'' \emph{IEEE Transactions on Intelligent Transportation Systems}, vol. 12, no. 4, pp. 1624-1639, 2018.

% KNN Graphs
\bibitem{dong2011efficient} W. Dong, M. Charikar, and K. Li, ``Efficient K-nearest neighbor graph construction for generic similarity measures,'' \emph{Proceedings of the 20th International Conference on World Wide Web}, pp. 577-586, 2011.

% Delaunay Triangulation
\bibitem{lee1980two} D. T. Lee and B. J. Schachter, ``Two algorithms for constructing a Delaunay triangulation,'' \emph{International Journal of Computer & Information Sciences}, vol. 9, no. 3, pp. 219-242, 1980.

% Gabriel Graphs
\bibitem{gabriel1969new} K. R. Gabriel and R. R. Sokal, ``A new statistical approach to geographic variation analysis,'' \emph{Systematic Biology}, vol. 18, no. 3, pp. 259-278, 1969.

% Relative Neighborhood Graphs
\bibitem{toussaint1980relative} G. T. Toussaint, ``The relative neighbourhood graph of a finite planar set,'' \emph{Pattern Recognition}, vol. 12, no. 4, pp. 261-268, 1980.

% Minimum Spanning Trees
\bibitem{graham1985euclidean} R. L. Graham and P. Hell, ``On the history of the minimum spanning tree problem,'' \emph{Annals of the History of Computing}, vol. 7, no. 1, pp. 43-57, 1985.

% β-Skeletons
\bibitem{kirkpatrick1985framework} D. G. Kirkpatrick and J. D. Radke, ``A framework for computational morphology,'' \emph{Computational Geometry}, pp. 217-248, 1985.

% Spectral Clustering
\bibitem{ng2001spectral} A. Y. Ng, M. I. Jordan, and Y. Weiss, ``On spectral clustering: Analysis and an algorithm,'' \emph{Advances in Neural Information Processing Systems}, vol. 14, pp. 849-856, 2001.

% Multi-view Anchor Graph-based Clustering
\bibitem{wang2019multi} Y. Wang, L. Wu, X. Lin, and J. Gao, ``Multiview spectral clustering via structured low-rank matrix factorization,'' \emph{IEEE Transactions on Neural Networks and Learning Systems}, vol. 29, no. 10, pp. 4833-4843, 2018.

% K-Means
\bibitem{arthur2007k} D. Arthur and S. Vassilvitskii, ``k-means++: The advantages of careful seeding,'' \emph{Proceedings of the Eighteenth Annual ACM-SIAM Symposium on Discrete Algorithms}, pp. 1027-1035, 2007.

% DBSCAN
\bibitem{ester1996density} M. Ester, H.-P. Kriegel, J. Sander, and X. Xu, ``A density-based algorithm for discovering clusters in large spatial databases with noise,'' \emph{Proceedings of the Second International Conference on Knowledge Discovery and Data Mining}, vol. 96, no. 34, pp. 226-231, 1996.

% Hierarchical Clustering
\bibitem{murtagh2012algorithms} F. Murtagh and P. Contreras, ``Algorithms for hierarchical clustering: an overview,'' \emph{Wiley Interdisciplinary Reviews: Data Mining and Knowledge Discovery}, vol. 2, no. 1, pp. 86-97, 2012.

% Affinity Propagation
\bibitem{frey2007clustering} B. J. Frey and D. Dueck, ``Clustering by passing messages between data points,'' \emph{Science}, vol. 315, no. 5814, pp. 972-976, 2007.

% Local Outlier Factor
\bibitem{breunig2000lof} M. M. Breunig, H.-P. Kriegel, R. T. Ng, and J. Sander, ``LOF: identifying density-based local outliers,'' \emph{Proceedings of the 2000 ACM SIGMOD International Conference on Management of Data}, pp. 93-104, 2000.

% Isolation Forest
\bibitem{liu2008isolation} F. T. Liu, K. M. Ting, and Z.-H. Zhou, ``Isolation forest,'' \emph{2008 Eighth IEEE International Conference on Data Mining}, pp. 413-422, 2008.

% One-Class SVM
\bibitem{scholkopf1999support} B. Schölkopf, R. C. Williamson, A. J. Smola, J. Shawe-Taylor, and J. C. Platt, ``Support vector method for novelty detection,'' \emph{Advances in Neural Information Processing Systems}, vol. 12, pp. 582-588, 1999.

% Statistical outlier detection
\bibitem{barnett1994outliers} V. Barnett and T. Lewis, ``Outliers in statistical data,'' \emph{Wiley Series in Probability and Statistics}, 1994.

% Dijkstra's Algorithm
\bibitem{dijkstra1959note} E. W. Dijkstra, ``A note on two problems in connexion with graphs,'' \emph{Numerische Mathematik}, vol. 1, no. 1, pp. 269-271, 1959.

% A* Algorithm
\bibitem{hart1968formal} P. E. Hart, N. J. Nilsson, and B. Raphael, ``A formal basis for the heuristic determination of minimum cost paths,'' \emph{IEEE Transactions on Systems Science and Cybernetics}, vol. 4, no. 2, pp. 100-107, 1968.

% Floyd-Warshall Algorithm
\bibitem{floyd1962algorithm} R. W. Floyd, ``Algorithm 97: shortest path,'' \emph{Communications of the ACM}, vol. 5, no. 6, p. 345, 1962.

% Bellman-Ford Algorithm
\bibitem{bellman1958routing} R. Bellman, ``On a routing problem,'' \emph{Quarterly of Applied Mathematics}, vol. 16, no. 1, pp. 87-90, 1958.

\end{thebibliography}

\end{document}
% that's all folks